Optimal (Multiway) Spatial Joins

### 2024-05-27 msj.tex
yufei

\extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$. Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $\bm{t} = (e_1, e_2, .., e_k)$ in the cartesian product $S_1 \times S_2 \times ... \times S_k$ as a $k$-dimensional vector such that $\bm{t}[i] = e_i$ for each $i \in [k]$. Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle. All logarithms have base 2 by default.

@@

In terms of techniques, our primary contribution is the revelation of a new property on the problem's mathematical structure. Fix any $k \ge 3$ and an {\em arbitrary} algorithm $\A$ for the $(k-1)$-SJ problem. For any $n \ge 1$, define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size at most $n$ and output size at most $\out$. Our core technical contribution is to establish:

@@

The theorem implies the existence of a recursive nature of $k$-SJ. Indeed, we will see that the $k$-SJ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all of which have input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + \out)$. Furthermore, if $\A$ is comparison-based, so is the $k$-SJ algorithm that ensues. For 2-SJ, we can set $\A$ simply to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$. Combining this with \eqref{eqn:main:reccurrence} yields a recurrence that relates the time complexity of $k$-SJ to that of $(k-1)$-SJ. Solving the recurrence yields:

@@

@@

>>>>

\extraspacing {\bf Math Conventions.} For any integer $x \ge 1$, we use $[x]$ to represent the set $\set{1, 2, ..., x}$. Given $k \ge 2$ sets $S_1, S_2, ..., S_k$ (of arbitrary elements), we often treat a $k$-tuple $(e_1, e_2, .., e_k)$ in the cartesian product $S_1 \times S_2 \times ... \times S_k$ as a $k$-dimensional vector $\bm{t}$ with $\bm{t}[i] = e_i$ for each $i \in [k]$. Every mention of the word ``rectangle'' henceforth will refer to an axis-parallel rectangle in $\real^2$. All logarithms have base 2 by default.

@@

In terms of techniques, our primary contribution is the revelation of a new property on the problem's mathematical structure. Fix any $k \ge 3$ and an {\em arbitrary} algorithm $\A$ for the $(k-1)$-SJ problem. Define function $F_{k-1}(n, \out)$ to return the worst-case running time of $\A$ on any instance of the $(k-1)$-SJ problem having input size at most $n$ and output size at most $\out$. Our core technical contribution is to establish:

@@

The theorem implies a recursive nature of $k$-SJ. Indeed, we will see that an $k$-SJ instance with input size $n$ and output size $\out$ can be converted to $O(k^3)$ instances of the $(k-1)$-SJ problem --- all having input size at most $n$ and output size at most $\out$ --- plus an additional processing cost of $O(n \log n + \out)$. Moreover, if $\A$ is comparison-based, so is the $k$-SJ algorithm that ensues. For 2-SJ, we can set $\A$  to the ``folklore algorithm'' mentioned in Section~\ref{sec:intro:prev}, which ensures $F_2(n, \out) = O(n \log n + \out)$. Combining this with \eqref{eqn:main:reccurrence} gives a recurrence that relates the time complexity of $k$-SJ to that of $(k-1)$-SJ. Solving the recurrence yields:

@@

\input{thm1.2}

@@

\input{hard-3d}
